The effects of ongoing context inference on motor learning

### 2021-11-11 adaptation/article/article.tex
Update on Overleaf.

\title{The effects of ongoing context inference on motor learning}
\shorttitle{Context-inference-dependent motor learning}
\author[1,2]{Cuevas Rivera, Darío}
\author[1,2]{Kiebel, Stefan J.}
\affil[1]{Chair of Neuroimaging, Faculty of Psychology. Technische Universität Dresden, 01062 Dresden, Germany.}
\affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
\affiliation{~}

@@

\abstract{Abstract}
Humans have been shown to adapt their movements when a sudden change to the dynamics of the environment is introduced, a phenomenon called motor adaptation. If the change is reverted, the adaptation is also quickly reverted. Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly. Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations. In this work, we use a model to explain the behavioral phenomena effected by imperfect contextual information. Specifically, we present a hierarchical computational model for motor adaptation based on exact Bayesian inference. This model explicitly takes into account contextual information and how the dynamics of context inference affect adaptation and action selection. We show how the proposed model provides a unifying explanation for four different experimentally established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of previously-learned adaptations on switching between tasks, in addition to (iii) the well-studied savings, de-adaptation and spontaneous recovery.

@@

We now briefly describe the three components of the model separately. For clarity, we start with the second component (i.e. motor adaptation), in order to introduce the terminology that we use throughout this work. For an overview, we show in \fref{fig:model} how these three components affect each other.

@@

Following these ideas, our model performs context inference that integrates several sources of information, including those typical for experimental studies such as visual cues \citep{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or the start of a new block of trials \citep{Ethier_Spontaneous_2008}, in addition to proprioceptive information, outcome prediction errors and reward prediction errors. As with motor adaptation, the context is inferred with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.

@@

We generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task. With this, we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.

We selected representative experimental studies that show savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues. Using our model, we showed how each of these effects can be explained by the specifics of context inference, which integrates all the available information (e.g. sensory cues, workspace location, reward and endpoint feedback), in some cases throughout many trials.

>>>>

\title{The effects of ongoing context inference on motor learning}
\shorttitle{Context-inference-dependent motor learning}
\author[1,2]{Cuevas Rivera, Darío}
\author[1,2]{Kiebel, Stefan J.}
\affil[1]{Chair of Neuroimaging, Faculty of Psychology, Technische Universität Dresden, 01062 Dresden, Germany.}
\affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
\affiliation{~}

@@

\abstract{Abstract}
Humans have been shown to adapt their movements when a sudden change to the dynamics of the environment is introduced, a phenomenon called motor adaptation. If the change is reverted, the adaptation is also quickly reverted. Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly. Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations. In this work, we introduce a computational model to explain the behavioral phenomena effected by imperfect contextual information. Specifically, we present a hierarchical model for motor adaptation based on exact Bayesian inference. This model explicitly takes into account contextual information and how the dynamics of context inference affect adaptation and action selection. We show how the proposed model provides a unifying explanation for four different experimentally established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of previously-learned adaptations on switching between tasks, in addition to (iii) the well-studied savings, de-adaptation and spontaneous recovery.

@@

We now briefly describe the three components of the model separately. For clarity, we start with the second component (i.e. motor adaptation), in order to introduce the terminology that we use throughout this work. For an overview, we show in \fref{fig:model} how these three components affect each other. The full mathematical description of the model can be found in the Methods.

@@

Following these ideas, our model performs context inference that integrates several sources of information, including those typical for experimental studies such as visual cues \citep{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or the start of a new block of trials \citep{Ethier_Spontaneous_2008}, in addition to proprioceptive information, outcome prediction errors and reward prediction errors. As with motor adaptation, the context is inferred as a probability distribution $q(\zeta)$ with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.

@@

We generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task. With this, we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had assumed the existence of multiple mechanisms.

We selected representative experimental studies that show the well-established effects of savings, quick de-adaptation, spontaneous recovery and the effects of sensory cues. Using our model, we showed how each of these effects can be explained by the specifics of context inference, which integrates all the available information (e.g. sensory cues, workspace location, reward and endpoint feedback), in some cases throughout many trials.
